---
title: "Reanalysis OSRI"
format: html
editor: visual
---

In this application, we perform a reanalysis of the data reported in De Roover & Vermunt (2019):

De Roover, K., & Vermunt, J. K. (2019). On the exploratory road to unraveling factor loading non-invariance: A new multigroup rotation approach. Structural Equation Modeling: A Multidisciplinary Journal, 26(6), 905-923.

# Data preparation

## Imports

```{r}
rm(list = ls())
```

## Set seed

```{r}
set.seed(17092024)
```

## Loading the data

Data available on OSF: <https://osf.io/rvhsw>

```{r}
OSRI <- read.csv("../data/OSRI_osfDeRoover.csv", sep = ";")
```

```{r}
table(OSRI$orientation)
```

The total sample size matches what is reported in De Roover & Vermunt (2019) but group sizes not. Upon checking if they erroneously used one of the questionnaire items as the indicator of sexual orientation, this appeared to be the case (it seems the scores on Q1 were used as the indicator of sexual orientation). This mistake probably also explains why their results deviate strongly from what can be expected based upon the design of the questionnaire. To make sure correct data are used, we worked on the original data availabe at <https://openpsychometrics.org/_rawdata/>.

## Loading the original data

```{r}
OSRI <- read.csv("../data/OSRI_original.csv", sep = ";")
head(OSRI)
```

## Select sample using criteria De Roover

```{r}
#Englisch speaking
OSRI_english <- OSRI[OSRI$engnat==1,]
#valid orientations
OSRI_english_valid <- OSRI_english[OSRI_english$orientation>0 & OSRI_english$orientation<6,]
#complete cases only
OSRI_clean <- OSRI_english_valid
for (j in c(1:44,54)){
  OSRI_clean <- OSRI_clean[OSRI_clean[,j]>0 & OSRI_clean[,j]<6,]
}
#first 2500
OSRI <- OSRI_clean[1:2500,]
#Select OSRI items + sexual orientation variable
OSRI <- OSRI[,c(1:44,54)]
```

```{r}
N <- table(OSRI$orientation) 
N
```

## Data wrangling

Group by orientation and standardize per group. Note: 1= Heterosexual, 2 = Bisexual, 3 = Homosexual, 4 = Asexual, 5 = Other.

```{r}
matrix1 <- as.matrix(OSRI[OSRI$orientation == 1, 1:44])
matrix1 <- scale(matrix1, center = TRUE, scale = TRUE)
matrix2 <- as.matrix(OSRI[OSRI$orientation == 2, 1:44])
matrix2 <- scale(matrix2, center = TRUE, scale = TRUE)
matrix3 <- as.matrix(OSRI[OSRI$orientation == 3, 1:44])
matrix3 <- scale(matrix3, center = TRUE, scale = TRUE)
matrix4 <- as.matrix(OSRI[OSRI$orientation == 4, 1:44])
matrix4 <- scale(matrix4, center = TRUE, scale = TRUE)
matrix5 <- as.matrix(OSRI[OSRI$orientation == 5, 1:44])
matrix5 <- scale(matrix5, center = TRUE, scale = TRUE)
groups <- 5

#DATA IN MG-REAFA FORMAT
DATA <- list(matrix1, matrix2, matrix3, matrix4, matrix5)

#DATA IN MPLUS/LSLX FORMAT
groupingvar <- rep(c(1:length(N)),rep(N))
DATA_tmp<-rbind(matrix1, matrix2, matrix3, matrix4, matrix5)
data_osri_lslx_mplus <- data.frame(cbind(DATA_tmp,groupingvar))
names(data_osri_lslx_mplus)[1:44] <- c("X1","X2","X3","X4","X5","X6","X7","X8","X9","X10",
                                 "X11","X12","X13","X14","X15","X16","X17","X18","X19","X20",
                                 "X21","X22","X23","X24","X25","X26","X27","X28","X29","X30",
                                 "X31","X32","X33","X34","X35","X36","X37","X38","X39","X40",
                                 "X41","X42","X43","X44")
names(data_osri_lslx_mplus)[45] <- "G"
```

We set the number of factors in correspondence with the design of the questionnaire (2 factors):

```{r}
factors <- 2
```

# Modelling: MG-REAFA

## Imports

```{r}
# load MASS
library(MASS)
```

```{r}
# load needed functions
source("../R/initial-loadings.R")
source("../R/multistart.R")
source("../R/pstr.R")
source("../R/joint-spca-finalizing.R")
source("../R/is-jspca.R")
source("../R/seqstrategy.R")
source("../R/performance-measures.R")
```

## Model selection

To determine the number of nonzero loadings and the strength of fusion of the loadings between the five groups, the sequential model selection strategy is used. Note that this takes a bit of time (a few minutes).

```{r}
model <- seqstrategy(DATA, R = factors, SP = 0.9, ALPH = 0.1, THR = 0.05, TOL = 2) # sequential strategy
```

We inspect the results as follows:

```{r}
model$details
model$cardinality
model$lambda
```

Note that the selected number of nonzero loadings is close to what can be expected as the questionnaire is designed to have 22 items that measure masculinity and 22 femininity.

## Stabilize solution

To save computing time, the sequential strategy uses a single starting value. Therefore, for the selected model, we perform an additional analysis that is based on multiple starting values. This analysis takes a few minutes.

```{r}
selmodel <- MULTISTART(DATA, R=factors, CARD = model$cardinality, LAMBDA = model$lambda, SP = seq(0.1, 0.9, by = 0.1), ALPH = seq(0.1, 0.9, by = 0.1), MAXITER = 20, EPS = 1e-4)
L <- selmodel$loadings
Loadings <- round(cbind(L[[1]], L[[2]], L[[3]], L[[4]], L[[5]]), 2)
colnames(Loadings) <- c("He1","He2","Bi1","Bi2","Ho1","Ho2","A1","A2","O1","O2")
Loadings
```

The loadings estimated by Regularized MGEAFA are approximately equal to the correlation of the observed variable with the estimated factor scores: the fusion penalty introduces shrinkage (over groups) and the cardinality constraint results in zero loadings. We illustrate this with the loadings of the variables on the first factor for the first group.

```{r}
rbind(cor(selmodel$scores[[1]][, 1], matrix1), selmodel$loadings[[1]][, 1])
```

## Inspect results

Here, we focus on inpecting the similarity of the loadings over groups. First, we create two matrices, one with the loadings on the first factor and one with those on the second factor.

```{r}
concL1 <- c()
concL2 <- c()
for (g in 1:groups) {
  concL1 <- cbind(concL1, L[[g]][, 1])
  concL2 <- cbind(concL2, L[[g]][, 2])
}
```

```{r}
concL1
concL2
```

To get an indication of the extent of measure noninvariance of each of the items, dissimilarity over the groups is calculated as the standard deviation of the loadings over groups (per factor and summed):

```{r}
round(sqrt(diag(var(t(concL1)))) + sqrt(diag(var(t(concL2)))), 3)
# Large(r) noninvariance: 24, 32, 31, 37,11, 13, 6, 
```

We also check if there are items that load (almost) 0 on both factors as this indicates that these should probably be removed from the questionnaire.

```{r}
index <- 1:44
index[rowSums(abs(cbind(concL1, concL2))) < 0.05]
# none
```

The following items have cross loadings and therefore deviate from the expected structure (each item loading only on one of the two factors):

```{r}
index[abs(concL1) > 0.05 & abs(concL2) > 0.05]
# items 8 11 12 19 20 21 25 31 33 37
```

## Save results

Write model selection results to csv:

```{r}
write.table(model$details, file = "OSRI_modelselection.txt", dec = ",", sep = "\t")
```

Write loadings to csv:

```{r}
write.table(Loadings, file = "OSRI_loadings_MGREAFA.txt", dec = ",", sep = "\t")
```

Save concatenaed loadings over groups and per factor:

```{r}
save(concL1, file = "OSRI_concl1.Rdata")
save(concL1, file = "OSRI_concl2.Rdata")
```

# Modelling: MPLUS AESEM

To run MPLUS, a recent version of mplus is needed. Furthermore, the MPLUSautomation package has to be installed.

```{r}
library(MplusAutomation)
varstr <- paste("usevariables = x1 - x44; GROUPING = G(1=group1 2=group2 3=group3 4=group4 5=group5);")
modelstr <- paste("f1-f2 BY x1-x44 (*1);")

m.penefa <- mplusObject(
  TITLE = "Exploratory Penalized Factor Analysis",
  VARIABLE = varstr,#"usevariables = v1 - v20; GROUPING = G(1=group1 2=group2);",
  ANALYSIS = "alignment = free; ROTATION = GEOMIN (ORTHOGONAL);",
  MODEL = modelstr,
  OUTPUT = "standardized;",#standardized loadings (instead of align) to compare with rmgeafa
  rdata = data_osri_lslx_mplus)

m.penefa.fit <- mplusModeler(m.penefa, dataout="osri.dat", 
                             modelout="osri.inp",run = TRUE, 
                             writeData = "ifmissing")

#POSTPROCESS MPLUS OUTPUT
loadingsparams <- m.penefa.fit$results$parameters$stdyx.standardized
loadings <- list()
card_mplus <- c()
for (g in 1:groups){
  loadings[[g]] <- matrix(nrow=44,ncol=factors)
  for (r in 1:factors){
    strR <- paste("F",r,".BY",sep="")
    strG <- paste("GROUP",g,sep="")
      index <- loadingsparams$paramHeader==strR & loadingsparams$Group==strG
      loadings[[g]][,r] <- as.numeric(loadingsparams$est[index])
#      if (dim(loadingsparams)[2]>4){
#        loadings[[g]][loadingsparams$pval[index] > 0.05,r] <- 0
#    }
  }
  card_mplus <- rbind(card_mplus,colSums(abs(loadings[[g]])>0.05))#count nr of zero loadings
}
mplusmodel <- list()
mplusmodel$loadings <- loadings
#permutation to align with rmgeafa: same for all groups!
Loadings_perm <- round(cbind(loadings[[1]][,c(2,1)], loadings[[2]][,c(2,1)], loadings[[3]][,c(2,1)],
                             loadings[[4]][,c(2,1)],loadings[[5]][,c(2,1)]), 2)

#Write loadings to csv
write.table(Loadings_perm, file = "OSRI_MPLUS.txt", dec = ",", sep = "\t")
```

# Modelling: LSLX

To run the code here below, the lslx package needs to be installed (available on CRAN).

WARNING: The following analysis may run very slowly (24 hours on a Dell Latitude 3420)

```{r}
library(lslx)

#specify multigroup efa model with penalized loadings
line1 <-paste("f1:~> ",paste("X",c(1:44),collapse = "+",sep = ""))
line2 <-paste("f2:~> ",paste("X",c(1:44),collapse = "+",sep = ""))
line3 <- paste("f1  <=>  f2")
line4 <-paste("f1  <=> fix(1)* f1")
line5 <-paste("f2  <=> fix(1)* f2")
model_mgfa <- paste(paste(line1,line2,line3,line4,line5,collapse = "\n", sep="\n"),collapse = "\n","\n")

#run lxlx analysis
start.time <- Sys.time()
lslx_mgfa <- lslx$new(model = model_mgfa,
                      data = data_osri_lslx_mplus,
                      group_variable = "G",
                      reference_group = "1")#,
lslx_mgfa$penalize_heterogeneity(
  block = c("y<-f","y<-1"),
  group = "2"
)
lslx_mgfa$penalize_heterogeneity(
  block = c("y<-f","y<-1"),
  group = "3"
)
lslx_mgfa$penalize_heterogeneity(
  block = c("y<-f","y<-1"),
  group = "4"
)
lslx_mgfa$penalize_heterogeneity(
  block = c("y<-f","y<-1"),
  group = "5"
)
lslx_mgfa$fit(penalty_method = "mcp", 
              lambda_grid = seq(.01, .60, .01), 
              delta_grid = c(1.5, 3.0, Inf))
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken #Time difference of 23 hours
save(lslx_mgfa,file="lslxOSRI.Rda")

load("lslxOSRI.Rda")
#lslx_mgfa$free_heterogeneity(block = "y <- f")
lslx_mgfa$summarize(selector = "bic", interval = FALSE)
lslxcoefs <- lslx_mgfa$test_coefficient(selector = "bic")

#POSTPROCESS results
loadings <- list()
pvals <- list()
card_lslx <- c()
for (g in 1:groups){
  loadings[[g]] <- matrix(nrow=44,ncol=factors)
  pvals[[g]] <- matrix(nrow=44,ncol=factors)
  for (r in 1:factors){
    strRG <- paste("<-f",r,"/",g,sep="")
    index <- grep(strRG,rownames(lslxcoefs))
    if (g==1){#reference group
      loadings[[g]][,r] <- lslxcoefs$estimate[index]
    } else {
      loadings[[g]][,r] <- lslxcoefs$estimate[index]+loadings[[1]][,r] 
    }
#    pvals[[g]][,r] <- lslxcoefs$p_value[index]
  }
  print(colSums(abs(loadings[[g]])<0.01))#count nr of zero loadings
  print(colSums(abs(loadings[[g]])<0.01 | pvals[[g]]>0.05))#count nr of zero loadings
    #card_lslx <- rbind(card_lslx,colSums(abs(loadings[[g]])>0.05 | pvals[[g]]<0.05))#count nr of zero loadings
  card_lslx <- rbind(card_lslx,colSums(abs(loadings[[g]])>0.05))#count nr of zero loadings
}

L_table_lslx <- round(cbind(loadings[[1]], loadings[[2]], loadings[[3]],
                      loadings[[4]], loadings[[5]]), 2)
L_table_lslx

#sign conversion+permutation for easier interpretation
L_table_lslx[,c(1,3,5,7,9)] <- -1*L_table_lslx[,c(1,3,5,7,9)]
L_table_lslx[,c(1:10)] <- L_table_lslx[,c(2,1,4,3,6,5,8,7,10,9)]

#Write loadings to csv
write.table(L_table_lslx, file = "OSRI_LSLX.txt", dec = ",", sep = "\t")
#fit indices
lslx_mgfa$extract_fit_index(selector="bic")

```
