---
title: "Didactical Example"
format: html
editor: visual
---

# Data Preparation

## Imports

```{r}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
rm(list = ls())
```

```{r}
library("MASS") # our functions use mvrnorm in the random initialization
```

## Creating the small data example

We rely on the simulation setup by De Roover et al. to create data for R=2 groups and G=3 components with a difference in the position of cross loadings between group 3 and groups 1 & 2 (groups 1 & 2 have same loadings):

```{r}
R <- 2 # 2 groups
G <- 3 # 3 groups
CL <- .2 #value cross loading
# P1
p1g1 <- c(rep(sqrt(.6), 4),rep(CL, 1), rep(0, 3))
p2g1 <- c( rep(CL,1), rep(0, 3),rep(sqrt(.6), 4))
P1 <- cbind(p1g1, p2g1)

# P2
P2 <- P1

# P3
p1g3 <- c(rep(sqrt(.6), 4), 0, 0, 0, rep(CL, 1))
p2g3 <- c(rep(0, 3), rep(CL, 1), rep(sqrt(.6), 4))
P3 <- cbind(p1g3, p2g3)

#all
cbind(P1,P2,P3)

#var-cov
PSI1 <- diag(1 - rowSums(P1^2))
SIGMA1 <- P1 %*% t(P1) + PSI1
PSI2 <- diag(1 - rowSums(P2^2))
SIGMA2 <- P2 %*% t(P2) + PSI2
PSI3 <- diag(1 - rowSums(P3^2))
SIGMA3 <- P3 %*% t(P3) + PSI3

#generate data
N <- 50
data1 <- mvrnorm(n = N, mu = rep(0, 8), Sigma = SIGMA1, empirical = TRUE)
data2 <- mvrnorm(n = N, mu = rep(0, 8), Sigma = SIGMA2, empirical = TRUE)
data3 <- mvrnorm(n = N, mu = rep(0, 8), Sigma = SIGMA3, empirical = TRUE)
DATA <- list(data1, data2, data3)
  
```

# Regularized MultiGroup Approximate EFA

## Imports

```{r}
source("../R/initial-loadings.R")
source("../R/multistart.R")
source("../R/pstr.R")
source("../R/joint-spca-finalizing.R")
source("../R/is-jspca.R")
source("../R/seqstrategy.R")
source("../R/performance-measures.R")
```

## Set seed

```{r}
set.seed(1808)
```

## Step 1: MODEL SELECTION, determine K and lambda

```{r}
modelsel <- seqstrategy(DATA, R = 2, SP = 0.9, ALPH = 0.1, THR = 0.05) # sequential strategy
```

Inspect results:

```{r}
modelsel$details
modelsel$cardinality # selected K
modelsel$lambda # selected lambda
```

## Step 2: OBTAIN ESTIMATES FOR SELECTED MODEL using multistart

```{r}
selmodel <- MULTISTART(DATA, R, CARD = modelsel$cardinality, LAMBDA = modelsel$lambda, SP = seq(0.1, 0.9, by = 0.1), ALPH = seq(0.1, 0.9, by = 0.1), MAXITER = 20, EPS = 1e-4)
L <- selmodel$loadings
L_table <- round(cbind(L[[1]], L[[2]], L[[3]]), 2)
L_table
```

## Save results

Write model selection results to csv file:

```{r}
write.table(modelsel$details, file = "SimpleExample_modelselection.txt", dec = ",", sep = "\t")
```

Write loadings to csv:

```{r}
write.table(L_table, file = "SimpleExampleQOL_loadings_RegMGEFA.txt", dec = ",", sep = "\t")
```

# MPLUS analysis

## Set seed

```{r}
library(MplusAutomation)
set.seed(510)
```

## Alignment analysis

```{r}
#Create data suitable for MPLUS
DATA <- cbind(data.frame(rbind(data1,data2,data3)),rep(c(1:G),rep(50,G)))
colnames(DATA) <- c("v1","v2","v3","v4","v5","v6","v7","v8","G")

#Create input file
varstr <- paste("usevariables = v1 - v8; GROUPING = G(1=group1 2=group2 3=group3);")
modelstr <- paste("f1-f2 BY v1-v8 (*1);")
m.penefa <- mplusObject(
  TITLE = "Exploratory Penalized Factor Analysis",
  VARIABLE = varstr,
  ANALYSIS = "alignment = free;",
  MODEL = modelstr,
  OUTPUT = "align;",
  rdata = DATA)

#MPLUS alignment analysis
m.penefa.fit <- mplusModeler(m.penefa, dataout="SimpleExample.dat", 
                             modelout="SimpleExample.inp",run = TRUE, 
                             writeData = "ifmissing")

#POSTPROCESS MPLUS OUTPUT
loadingsparams <- m.penefa.fit$results$parameters$unstandardized
loadings <- list()
for (g in 1:G){
  loadings[[g]] <- P1
  for (r in 1:R){
    strR <- paste("F",r,".BY",sep="")
    strG <- paste("GROUP",g,sep="")
    index <- loadingsparams$paramHeader==strR & loadingsparams$Group==strG
    loadings[[g]][,r] <- as.numeric(loadingsparams$est[index])
  }
}

```

Write output to csv:

```{r}
write.table(loadingsparams, file = "SimpleExampleMPLUS_loadingparams.txt", dec = ",", sep = "\t")
L_table_mplus <- round(cbind(loadings[[1]], loadings[[2]], loadings[[3]]), 2)
L_table_mplus
write.table(L_table_mplus, file = "SimpleExampleMPLUS_loadings.txt", dec = ",", sep = "\t")
```

## Regularization path

### LSLX analysis

```{r}
library(lslx)
set.seed(1810)
```

### Regularization

```{r}
data_lslx <- data.frame(rbind(data1,data2,data3))
grouplabel <- as.factor(rep(c("G1","G2","G3"),c(rep(50,G))))
data_lslx <- data.frame(cbind(data_lslx,grouplabel))
samplex <- seq(1:8)
line3 <-paste("f1:~> ",paste("X",samplex[1:8],collapse = "+",sep = ""))
line4 <-paste("f2:~> ",paste("X",samplex[1:8],collapse = "+",sep = ""))
linex <- paste("f1  <=>  f2")
line5 <-paste("f1  <=> fix(1)* f1")
line6 <-paste("f2  <=> fix(1)* f2")
model_mgfa <- paste(paste(line3,line4,linex,line5,line6,collapse = "\n", sep="\n"),collapse = "\n","\n")

#run lxlx analysis: NO alignment
lslx_mgfa <- lslx$new(model = model_mgfa,
                      data = data_lslx,
                      group_variable = "grouplabel")#,
lslx_mgfa$fit(penalty_method = "mcp", 
              lambda_grid = seq(.01, .60, .01), 
              delta_grid = c(1.5, 3.0, Inf))
lslx_mgfa$summarize(selector = "bic", interval = FALSE)
lslxcoefs <- lslx_mgfa$extract_coefficient(selector = "bic")
lslx_mgfa$test_coefficient(selector="bic")

#run lxlx analysis: Alignment to GROUP1
lslx_mgfa <- lslx$new(model = model_mgfa,
                      data = data_lslx,
                      group_variable = "grouplabel",
                      reference_group = "G1")#,
lslx_mgfa$penalize_heterogeneity(
  block = c("y<-f","y<-1"),
  group = "G2"
)
lslx_mgfa$penalize_heterogeneity(
  block = c("y<-f","y<-1"),
  group = "G3"
)
lslx_mgfa$fit(penalty_method = "mcp", 
              lambda_grid = seq(.01, .60, .01), 
              delta_grid = c(1.5, 3.0, Inf))
lslx_mgfa$summarize(selector = "bic", interval = FALSE, include_faulty = T)
lslxcoefs <- lslx_mgfa$extract_coefficient(selector = "bic", include_faulty = T)

error_count <- 0

result <- tryCatch(
  {
    lslx_mgfa$summarize(selector = "bic")   # or whichever call triggers the error
  },
  error = function(e) {
    if (grepl("PL estimate under EACH penalty level.*nonconverged result",
              conditionMessage(e),
              ignore.case = TRUE)) {
      error_count <<- error_count + 1
      message("Detected the nonconverged-all-lambda error.")
      return(NULL)
    } else {
      stop(e)  # rethrow other errors
    }
  }
)

error_count
lslx_mgfa$test_coefficient(selector="bic", include_faulty = T)
```

```{r}
loadings <- list()
for (g in 1:G){
  loadings[[g]] <- P1
  for (r in 1:R){
    strRG <- paste("<-f",r,"/",levels(grouplabel)[g],sep="")
    index <- grep(strRG,names(lslxcoefs))
    if (g==1){#reference group
      loadings[[g]][,r] <- lslxcoefs[index]
    } else {
      loadings[[g]][,r] <- lslxcoefs[index]+loadings[[1]][,r] 
    }
  }
}
L_table_lslx <- round(cbind(loadings[[1]], loadings[[2]], loadings[[3]]), 2)
L_table_lslx
write.table(L_table_lslx, file = "SimpleExampleLSLX_loadings.txt", dec = ",", sep = "\t")
```
